Global phase diagram and possible quantum spin liquid in the triangular $J-K-\Gamma$ model

### 2019-12-31 paper/paper.tex
20191231 12:18:40

\author{Shi Wang}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}
\author{Shun-Li Yu}
\email{slyu@nju.edu.cn}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}
\author{Jian-Xin Li}
\email{jxli@nju.edu.cn}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}

@@

\begin{abstract}
    The global ground state phase diagram of the Heisenberg-Kitaev-$\Gamma$ model on a triangular lattice is studied using classical Monte Carlo and exact diagonalization. \textcolor{red}{$\cdots \cdots$}
\end{abstract}

@@

\subsection{Exact Diagonalization}
We perform a Lanczos ED calculation of the GS energy of the Hamiltonian \eqref{eq:Hamiltonian} on several $4 \times 6$ clusters with periodic boundary conditions (PBC). To detect quantum phase transitions, the second derivatives of the ground-state energy, $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ were computed and looked for singular features that indicate changes in the groun-state characteristics. Phases containing exactly solvable or well-understood points, such as the FM, 120$^\circ$ N\'{e}el, stripe and Dual N\'{e}el can be readily identified. The remaining phases were identified mainly by examining the static magnetic structure factor (SMSF)
\begin{equation}
    S(\mathbf{Q}) = \frac{1}{N} \sum_{ij}e^{i\mathbf{Q}\cdot(\mathbf{R}_i - \mathbf{R}_j)} \langle \mathbf{S}_i \cdot \mathbf{S}_j \rangle
    \label{eq:StaticStructureFactor}
\end{equation}
where N is the total number of spins, $\mathbf{R}_i$ is the position of site $i$ and $\langle \mathbf{S}_i \cdot \mathbf{S}_j \rangle$ is the average over the ground state.

@@

\subsection{Moment direction of the classcial FM order}
For the classical FM order, the energy of the $J-K-\Gamma$ model per lattice site is given by
\begin{equation}
    E_{FM}^{c} = (3J + K) + 2\Gamma(S^xS^y + S^yS^z + S^zS^x)
    \label{eq:EcFM}
\end{equation}
On this level, the moment direction of the classical FM state is determined solely by $\Gamma$ and the problem becomes finding the global minimum and maximum of the multi-variable function $f(S^x, S^y, S^z) = S^xS^y + S^yS^z + S^zS^x$ with the constraint $|\mathbf{S}| = 1$. $f(S^x, S^y, S^z)$ takes maximum value $f_{max}=1$ at $S^x=S^y=S^z=\pm 1/\sqrt{3}$ and minimum value $f_{min}=-0.5$ when the conditions $S^x + S^y + S^z = 0$ and $|\mathbf{S}| = 1$ are fulfilled. That is to say, for $\Gamma > 0$ the moment of the classcial FM state prefers to lie in the plane that is perpendicular to the $[111]$ direction and when $\Gamma$ is negative, the moment would along the $[111]$ or $[\bar{1}\bar{1}\bar{1}]$ direction.

@@

In the case of stripe order, there are three degenerate spin configurations as illustrated in Fig. \ref{fig:PhaseDiagram}(b)-(d). For the spin configuration shown in Fig. \ref{fig:PhaseDiagram}(b), the energy of the model Hamiltonian \eqref{eq:Hamiltonian} is
\begin{IEEEeqnarray}{rCl}
    E_{StripeX}^{c} & = & -(J + K) + 2KS^xS^x \nonumber \\
        && + 2\Gamma(S^yS^z - S^xS^y - S^zS^x)
    \label{eq:EcStripeX}
\end{IEEEeqnarray}
and the energy for spin configuration in Fig. \ref{fig:PhaseDiagram}(c) and \ref{fig:PhaseDiagram}(d) can be obtained by a cyclic permutation $x \rightarrow y \rightarrow z \rightarrow x$. The moment direction of the classical stripe order is determined by the anisotropy parameters $K$ and $\Gamma$. In general, $E_{StripeX}^{c}$ has three extreme points where the first derivatives equal zero
\begin{align}
    &\mathbf{S}_0: \quad S_{0}^{y}=-S_{0}^{z} = \pm1/\sqrt{2}, \quad S_{0}^{x} = 0\\
    &\mathbf{S}_1: \quad S_{1}^{y}=S_{1}^{z} = f_{1}(K, \Gamma), \quad S_{1}^{x} = g_{1}(K, \Gamma) S_{1}^{y}\\
    &\mathbf{S}_2: \quad S_{2}^{y}=S_{2}^{z} = f_{2}(K, \Gamma), \quad S_{2}^{x} = g_{2}(K, \Gamma) S_{2}^{y}
\end{align}
see Appendix A for their explicit expressions.

For $\Gamma=0, K>0$, the stripe order shown in Fig. \ref{fig:PhaseDiagram}(b) prefers to lie in the $yz$ plane. A definite positive $\Gamma$ would fix the ordered moment to the $\mathbf{S}_0$ direction and negative $\Gamma$ drives it to the $\mathbf{S}_1$ direction. When $\Gamma=0, K<0$, the moment would along the $x$ axis to have lowest energy and both positive and negative $\Gamma$ make it deviate from the $x$ axis and point the $\mathbf{S}_1$ direction.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{(Color online) Map of the probabilities.}
     \label{fig:Proabilities}
\end{figure}

@@

>>>>

\author{Shi Wang}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}

\author{\textcolor{cyan}{XXX}}
\affiliation{\textcolor{cyan}{XXX}}
\author{\textcolor{cyan}{XXX}}
\affiliation{\textcolor{cyan}{XXX}}

\author{Shun-Li Yu}
\email{slyu@nju.edu.cn}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}
\author{Jian-Xin Li}
\email{jxli@nju.edu.cn}
\affiliation{National Laboratory of Solid State Microstructures and School of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}

@@

\begin{abstract}
\textcolor{red}{The global ground state phase diagram of the Heisenberg-Kitaev-$\Gamma$ model on a triangular lattice is studied using classical Monte Carlo and exact diagonalization. $\cdots \cdots$}
\end{abstract}

@@

\subsection{Exact Diagonalization}
We perform a Lanczos ED calculation of the GS energy of the Hamiltonian \eqref{eq:Hamiltonian} on several $4 \times 6$ clusters with periodic boundary conditions (PBC). To detect quantum phase transitions, the second derivatives of the ground-state energy, $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ were computed and looked for singular features that indicate changes in the ground-state characteristics. Phases containing exactly solvable or well-understood points, such as the FM, 120$^\circ$ N\'{e}el, stripe and Dual N\'{e}el can be readily identified. The remaining phases were identified mainly by examining the static magnetic structure factor (SMSF)
\begin{equation}
    S(\mathbf{Q}) = \frac{1}{N} \sum_{ij}e^{i\mathbf{Q}\cdot(\mathbf{R}_i - \mathbf{R}_j)} \langle \mathbf{S}_i \cdot \mathbf{S}_j \rangle
    \label{eq:StaticStructureFactor}
\end{equation}
where N is the total number of spins, $\mathbf{R}_i$ is the position of site $i$ and $\langle \mathbf{S}_i \cdot \mathbf{S}_j \rangle$ is the average over the ground state.

@@

\subsection{Moment direction of the classcial FM order}
For the classical FM order, the energy of the $J-K-\Gamma$ model per lattice site is given by
\begin{equation}
    E_{FM}^{c} = (3J + K) + 2\Gamma(S^xS^y + S^yS^z + S^zS^x)
    \label{eq:EcFM}
\end{equation}
On this level, the moment direction of the classical FM state is determined solely by $\Gamma$ and the problem becomes finding the global minimum and maximum of the multi-variable function $f(S^x, S^y, S^z) = S^xS^y + S^yS^z + S^zS^x$ with the constraint $|\mathbf{S}| = 1$. $f(S^x, S^y, S^z)$ takes maximum value $f_{max}=1$ at $S^x=S^y=S^z=\pm 1/\sqrt{3}$ and minimum value $f_{min}=-0.5$ when the conditions $S^x + S^y + S^z = 0$ and $|\mathbf{S}| = 1$ are fulfilled. The condition $S^x + S^y + S^z = 0$ specify a plane that is perpendicular to the $[111]$ direction, and considering the reference frame shown in Fig. \ref{fig:ModelDefinition}(c) this is actually the plane of the triangular lattice. That is to say, for $\Gamma > 0$ the moment of the classcial FM state prefers to lie in the lattice plane and when $\Gamma$ is negative, the moment would perpendicular to the lattice plane.

Our classical Monte Carlo simulation found FM ordered state with the right moment direction for majority of the area marked as FM phase (see Fig. 2 in Ref \onlinecite{PhysRevB.92.165108}) except for the pure $\Gamma$ term. For pure positive $\Gamma$ interaction, apart from the FM order lie in the lattice plane, we also found disordered states which have the same energy as the FM state but no states with lower energy were found. When $\Gamma=-1$, in addition to the FM order perpendicular to the lattice plane, other long-range ordered states which are energetically degenerate with the FM state appear (see Appendix A for details). Based on these observation, it is reasonable to say that the ground state of the classical $\Gamma$ model is highly degenerate though the degeneracy may be lifted by introducing other interactions. For example, a ferromagnetic ($J<0$) Heisenberg interaction would break the degeneracy and select the FM order as the ground state which was verified by our Monte Carlo simulation. On the other hand, when both positive Heisenberg and Kitaev interactions are included, which introduces further frustation to the system, the FM order is unlikely to be selected. For $\Gamma < 0$, the spins form stripe order and for positive $\Gamma$ the system remains disordered.

@@

In the case of stripe order, there are three degenerate spin configurations as illustrated in Fig. \ref{fig:PhaseDiagram}(b)-(d). For the spin configuration shown in Fig. \ref{fig:PhaseDiagram}(b), the energy of the model Hamiltonian \eqref{eq:Hamiltonian} is
\begin{IEEEeqnarray}{rCl}
    E_{StripeX}^{c} & = & -(J + K) + 2 K S^x S^x \nonumber \\
        && \negmedspace {} + 2 \Gamma (S^y S^z - S^z S^x - S^x S^y)
    \label{eq:EcStripeX}
\end{IEEEeqnarray}
and the energy for spin configuration in Fig. \ref{fig:PhaseDiagram}(c) and \ref{fig:PhaseDiagram}(d) can be obtained by a cyclic permutation $x \rightarrow y \rightarrow z \rightarrow x$. The moment direction of the classical stripe order is determined by the anisotropy parameters $K$ and $\Gamma$. In general, $E_{StripeX}^{c}$ has three extreme points where the first derivatives equal zero
\begin{align}
    &\mathbf{S}_0: \quad S_{0}^{y}=-S_{0}^{z} = \pm1/\sqrt{2}, \quad S_{0}^{x} = 0\\
    &\mathbf{S}_1: \quad S_{1}^{y}=S_{1}^{z} = f_{1}(K, \Gamma), \quad S_{1}^{x} = g_{1}(K, \Gamma) S_{1}^{y}\\
    &\mathbf{S}_2: \quad S_{2}^{y}=S_{2}^{z} = f_{2}(K, \Gamma), \quad S_{2}^{x} = g_{2}(K, \Gamma) S_{2}^{y}
\end{align}
see Appendix A for their explicit expressions.

For $\Gamma=0, K>0$, it can be clearly seen from Eq\eqref{eq:EcStripeX} that the stripe order shown in Fig. \ref{fig:PhaseDiagram}(b) prefers to lie in the $yz$ plane. A definite positive $\Gamma$ would fix the ordered moment to the $\mathbf{S}_0$ direction whereas negative $\Gamma$ drives it to the $\mathbf{S}_1$ direction. When $\Gamma=0, K<0$, the moment would along the $x$ axis to have lowest energy, both positive and negative $\Gamma$ make it deviate from the $x$ axis and point the $\mathbf{S}_1$ direction.

@@

@@

\section{Moment Direction-Exact Diagonalization Results}
\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{(Color online) Map of the probabilities.}
     \label{fig:Proabilities}
\end{figure}

\textcolor{red}{Exact diagonalization results of the moment direction for FM and stripe phases. Left blank currently.}
